Polynomial phase estimation by least squares phase unwrapping

### 2013-07-01 paper.tex
added to plots and paper

Tables~\ref{tablm3} and~\ref{tablm5} \dots BLERG

\begin{table}[htbp]
\centering
\begin{tabular}{lrrrrr}
 & LSU & CPF & PHAF & HAF & ZW \\ \toprule
$N=49$ &  &  &  &  &  \\ 
$N=99$ & 6 & 3 & 5 & 7 & 27 \\ 
$N=199$ & 3 & 1 & 3 & 5 & 23 \\ 
$N=299$ & 2 & -1 &  & 4 & \\ 
$N\to\infty$ & $-\infty$ & ? & ? & 3.77  & ? \\ \bottomrule
\end{tabular}
\caption{Estimated SNR threshold in decibels with $m=3$}
\label{tablm3}
\end{table}

\begin{table}[htbp]
\centering
\begin{tabular}{lrrrrr}
 & LSU & CPF-HAF & PHAF & HAF & ZW \\ \toprule
$N=49$ &  &  &  & &  \\ 
$N=99$ &  15 &  12 & 13 &  19 & $>100$ \\ 
$N=199$ & 12 & 10 & 12 & 18 & 103 \\ 
$N=299$ &  &  &  & &  \\ 
$N\to\infty$ & $-\infty$ & $\approx 6.5$ & ?  & 16.6 & ? \\ \bottomrule
\end{tabular}
\caption{Estimated SNR threshold in decibels with $m=5$}
\label{tablm5}
\end{table}

Figure~\ref{plot:bench} shows the running time of the LSU estimator computed using the $K$-best method versus the HAF estimator for polynomial phase signals of order 3.  The HAF estimator is computed using the fast Fourier transform and has complexity $O(N \log N)$.   Wtih $K=20N$ the LSU estimator computed using the $K$-best method requires $O(N^3 \log N)$ operations in the worst case.  The $K$-best method tends to be faster when the SNR is large.  This is due to the algorithm rapidly finding a nearest point, and subsequently terminating branches in its search tree early.  On the other hand, when the SNR is small, all branches in the tree tend to be searched.  Interestingly the LSU estimator computed using the $K$-best algorithm is comparable with the HAF when the SNR is large and when $N$ is approximately less than 100.  The LSU estimator can be computed in a reasonable amount of time (a few minutes) when $N=1000$.  The algorithm becomes prohibitively expensive for large $N$.  Benchmarks have been performed with polynomial phase signals of order larger than $3$.  The results are omitted due to space constraints.  The running time of the LSU estimator does not appear to change significantly as $m$ increases.  This is expected since a nearest point in a lattice of approximately the same dimension $N-m-1$ must be computed.  The complexity of the HAF estimator grows linearly with $m$.

>>>>

Tables~\ref{tablm1},~\ref{tablm3} and~\ref{tablm5} show the values of SNR at which the estimators reach a threshold for polynomial phase signals of order $m=1,3$ and $5$.  The values in the tables are approximated using Monte-Carlo simulation with $X_1, \dots, X_N$ pseudorandomly generated independent and identically distributed circularly symmetric complex Gaussian.  When $m=3$ the true coefficients are as in Figure~\ref{plot:m3small} and when $m=5$ the true coefficients are as in Figure~\ref{plot:m5} so that the coefficients are inside the range acceptable for the HAF, CPF and CPF-HAF estimators.  When $m=1$ (single frequency estimation) the LSU estimator can be computed exactly using the algorithm from~\cite{McKilliamFrequencyEstimationByPhaseUnwrapping2009}.  In this case the value at which the LSU estimator reaches a threshold is similar to the maximum likelihood periodogram estimator~\cite{Quinn2001,Hannan1973,Rife1974}.  This property was observed in~\cite{McKilliamFrequencyEstimationByPhaseUnwrapping2009}.  Both the LSU and periodogram estimators are strongly consistent, and so, the threshold values can be made arbitrarily small by choosing $N$ large.  This is indicated in the bottom row of Table~\ref{tablm1} by the $-\infty$ symbols.  When $m>1$ the LSU estimator can be exactly computed using the sphere decoder~\cite{Agrell2002,Viterbo_sphere_decoder_1999} in a reasonable amount of time when $N < 60$.  When $N > 60$ we use the approximate $K$-best algorithm.  We observe that the $K$-best algorithm produces estimates with the same threshold as the exactly computed LSU estimator when $N = 19$ and $49$.  As $N\rightarrow\infty$ the asymptotic threshold of the HAF estimator has been studied in~\cite{Porat_asympt_HAF_DPT_1996} and these values are indicated in the bottom rows of tables~\ref{tablm3} and~\ref{tablm5}.  By Theorem~\ref{thm:asymp_proof} the LSU estimator is strongly consistent and the threshold can be made as small as desired by choosing $N$ large.  Whether the LSU estimator computed approximately using the $K$-best algorithm is strongly consistent is an open theoretical question.  That the $K$-best algorithm provides a close approximation of the exact LSU estimator for small $N$ is an encouraging observation.

\begin{table}[t]
\centering
\begin{tabular}{lrr}
 & LSU & ML  \\ \toprule
$N=16$ & 5 & 2  \\ 
$N=64$ & -2 & -4 \\ 
$N=256$ & -6 & -8 \\ 
$N=1024$ & -14 & -14 \\ 
$N\to\infty$ & $-\infty$ & $-\infty$  \\ \bottomrule
\end{tabular}
\caption{Estimated SNR threshold in decibels with $m=1$}
\label{tablm1}
\end{table}

\begin{table}[t]
\centering
\begin{tabular}{lrrrrr}
 & LSU & LSU ($K$-best) & CPF & HAF & ZW \\ \toprule
$N=19$ & 10 & 10 & 8 & 11 & 31 \\ 
$N=49$ & 7 & 7 & 5 & 9 & 29 \\ 
$N=99$ & & 6 & 3 & 5 & 27 \\ 
$N=199$ & & 3 & 1 & 3 & 23 \\ 
$N=299$ & & 2 & -1 &  & \\ 
$N\to\infty$ & $-\infty$ & & & 3.77 & \\ \bottomrule
\end{tabular}
\caption{Estimated SNR threshold in decibels with $m=3$}
\label{tablm3}
\end{table}

\begin{table}[t]
\centering
\begin{tabular}{lrrrrr}
 & LSU & LSU ($K$-best) & CPF-HAF & HAF & ZW \\ \toprule
$N=19$ & 22  & 22 & 17 & 23 & $>100$ \\ 
$N=49$ & 17 & 17 & 14 & 21 & $>100$  \\ 
$N=99$ &  & 15 &  12 &  19 & $>100$ \\ 
$N=199$ & & 12 & 10 & 18 & 103 \\ 
$N=299$ &  &  &  & &  \\ 
$N\to\infty$ & $-\infty$ &  &  & 16.6 &  \\ \bottomrule
\end{tabular}
\caption{Estimated SNR threshold in decibels with $m=5$}
\label{tablm5}
\end{table}

Figure~\ref{plot:bench} shows the running time of the LSU estimator computed using the $K$-best method versus the HAF estimator for polynomial phase signals of order 3.  The HAF estimator is computed using the fast Fourier transform and has complexity $O(N \log N)$.   With $K=20N$ the LSU estimator computed using the $K$-best method requires $O(N^3 \log N)$ operations in the worst case.  The $K$-best method tends to be faster when the SNR is large.  This is due to the algorithm rapidly finding a nearest point, and subsequently terminating branches in its search tree early.  On the other hand, when the SNR is small, all branches in the tree tend to be searched.  Interestingly the LSU estimator computed using the $K$-best algorithm is comparable with the HAF when the SNR is large and when $N$ is approximately less than 100.  The LSU estimator can be computed in a reasonable amount of time (a few minutes) when $N=1000$.  The algorithm becomes prohibitively expensive for large $N$.  Benchmarks have been performed with polynomial phase signals of order larger than $3$.  The results are omitted due to space constraints.  The running time of the LSU estimator does not appear to change significantly as $m$ increases.  This is expected since a nearest point in a lattice of approximately the same dimension $N-m-1$ must be computed.  The complexity of the HAF estimator grows linearly with $m$.
